A Convolutional Neural Network for Outsole Recognition\tnoteref{t1}

### 2019-06-06 paper/paper.Rnw
Changing methods tone

%\svp{VGG16/ImageNet specifically. Note that the structure of the model is distinct from the fitted weights.}
%\svp{Transfer Learning - using CNNs with less data. Make the point that this is what makes our method different from some of the others we've already cited - training a new classifier allows us to work within already-existing semantic contexts that examiners use.}
%\mt{Now that I can see the whole dang methods section, I think I prefer the order of the methods section as it is now, rather than the order suggested by the above comments, unless you disagree}

@@

Many convolutional neural networks have been trained on a set of images known as ImageNet; these images are hierarchically labeled, so the same image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity \citep{dengImageNetLargeScaleHierarchical}. CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously. As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks. \autoref{fig:vgg16layers} shows a selection of filters from VGG16, where each filter detects the presence of specific shapes and colors. Filters in earlier convolutional blocks detect simple features, such as colors and blobs, while later filters detect more complex combinations of features. Each of the five convolutional blocks has between 128 and 1,536 filters, and a single image is made up of some combination of the thousands of features that are detected by those filters.

\fix{Transition needed?: The structure of the model is distinct from the fitted weights.} Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated \citep{oquabLearningTransferringMidlevel2014}. In many cases, the entire model base is used, and fitted with a new model head; the modularity of CNNs makes this process relatively simple, and allows researchers to leverage pre-trained networks when working with diffierent sets of image data. Transfer learning allows CNNs to be applied to smaller datasets of several thousand images, reduces the amount of computational time required to fit the model, and provides boosted performance compared to training a new model from scratch \citep{yosinskiHowTransferableAre2014}. It has been successfully applied in automatic classification of medical images \citep{shinDeepConvolutionalNeural2016} as well as in various applications of shoe forensics \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017}.

@@

\svp{Moving towards ... how did we train the model}
\paragraph{Annotated Training Data}

@@

\svp{It might be good to add in some (new) statistics relating brand information to common geometric features in tread - show, for instance, the top N brands with bowties vs. the top N brands with polygons. }

@@

Labeled images are scarce relative to the amount of data necessary to train a neural network. One solution to this scarcity is to artificially enlarge the data set using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}. Augmentation is the transformation of original input data using image operations such as cropping, zoom, skew, rotation, and color balance modification in order to distort or alter the image while maintaining the essential features corresponding to the label. This process reduces the potential for overfitting the model to the specific set of image data used during the training process, and also increases the amount of data available for training; examples of pre- and post-augmentation images are shown in \autoref{fig:augmented}.

@@

The \Sexpr{nrow(dfunion)} images were split such that 60\% were used for training. Since the categories do not exist in equal proportion in the labeled data, the training data were weighted by proportion during the training process to prevent the loss function from being overwhelmed by more frequent categories. Of the remaining 40\% of data, half were used for validation, to monitor the training process, and the remaining data were for testing the performance of the fitted model.

@@

\autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process. Overfitting, or fitting a model which performs too well on the training data relative to the validation data, is seen when the validation loss starts to increase after reaching a global minimum. Alternately, underfitting occurs when the validation accuracy is still increasing when model optimization is terminated. Neither of these outcomes appears in \autoref{fig:training-accuracy}, indicating that the model optimization process was halted at an appropriate epoch.

\svp{Starting into How CoNNOR performs}
\subsection{Model Accuracy}

>>>>

%\svp{VGG16/ImageNet specifically. Note that the structure of the model is distinct from the fitted weights.}
%\svp{Transfer Learning - using CNNs with less data. Make the point that this is what makes our method different from some of the others we've already cited - training a new classifier allows us to work within already-existing semantic contexts that examiners use.}

%\mt{Now that I can see the whole dang methods section, I think I prefer the order of the methods section as it is now, rather than the order suggested by the above comments, unless you feel strongly otherwise.}

@@

Many convolutional neural networks have been trained on a set of images known as ImageNet; these images are hierarchically labeled, so the same image of a dog may have labels of sky, grass, and trees, and the dog may be labeled as ``mammal'', ``dog'', ``retriever'', and ``golden retriever'' at increasing levels of label complexity \citep{dengImageNetLargeScaleHierarchical}. CNNs trained on ImageNet are optimized for general human-like vision, that is, the ability to recognize a large set of different features simultaneously. As a result, the base of networks trained on these sets are often used for transfer learning, because the initial layers are broadly generalizable to a wide variety of more specific image labeling tasks. \autoref{fig:vgg16layers} shows a selection of filters from VGG16, where each filter detects the presence of specific shapes and colors. \mt{Each row of filters in \autoref{fig:vgg16layers} corresponds to one of the five convolutional blocks of VGG16.} Filters in earlier convolutional blocks detect simple features, such as colors and blobs, while later filters detect more complex combinations of features. Each of the five convolutional blocks has between 128 and 1,536 filters, and a single image is made up of some combination of the thousands of features that are detected by those filters.

%\fix{Transition needed?: The structure of the model is distinct from the fitted weights.}
\mt{Each layer in the model structure is comprised of weights and connections between weights that have been optimized during training.} Transfer learning in neural network contexts involves freezing the weights of the pre-trained model base taken from a network trained on a more general set of images, so that during the training process, only the weights of the newly-added classifier are updated \citep{oquabLearningTransferringMidlevel2014}. In many cases, the entire model base is used, and fitted with a new model head; the modularity of CNNs makes this process relatively simple, and allows researchers to leverage pre-trained networks when working with diffierent sets of image data. Transfer learning allows CNNs to be applied to smaller datasets of several thousand images, reduces the amount of computational time required to fit the model, and provides boosted performance compared to training a new model from scratch \citep{yosinskiHowTransferableAre2014}. It has been successfully applied in automatic classification of medical images \citep{shinDeepConvolutionalNeural2016} as well as in various applications of shoe forensics \citep{kongCrossDomainForensicShoeprint2017,kongCrossDomainImageMatching2019a,zhangAdaptingConvolutionalNeural2017}.

@@

%\svp{Moving towards ... how did we train the model}
\paragraph{Annotated Training Data}

@@

%\svp{It might be good to add in some (new) statistics relating brand information to common geometric features in tread - show, for instance, the top N brands with bowties vs. the top N brands with polygons. }

\mt{The brand plots are still displaying weirdly for me, and all the figures are really squished and cut off?}

@@

%Labeled images are scarce relative to the amount of data necessary to train a neural network.
\mt{Although transfer learning substantially reduces the amount of data required to train a CNN, l}abeled images are scarce \fix{[difficult to generate]} relative to the amount of data necessary to train a neural network \mt{with adequate performance}. \fix{} One solution to this scarcity is to artificially enlarge the data set using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}. Augmentation is the transformation of original input data using image operations such as cropping, zoom, skew, rotation, and color balance modification in order to distort or alter the image while maintaining the essential features corresponding to the label. This process reduces the potential for overfitting the model to the specific set of image data used during the training process, and also increases the amount of data available for training; examples of pre- and post-augmentation images are shown in \autoref{fig:augmented}.

@@

The \Sexpr{nrow(dfunion)} images were split such that 60\% were used for training. Since \mt{some categories are much more common than others,}
%the categories do not exist in equal proportion in the labeled data,
the training data were weighted by proportion during the training process to prevent the \mt{model learning} from being overwhelmed by more frequent categories. Of the remaining 40\% of data, half were used for validation, to monitor the training process, and the remaining data were for testing the performance of the fitted model.

@@

\autoref{fig:training-accuracy} shows the training and validation accuracy and loss at each epoch of the fitting process. Overfitting \mt{occurs when a model learns the training data so well that its understanding of the categories becomes specific to only the training cases, which, in turn, leads to poor prediction of new images that were not in the training set.}
%Overfitting, or fitting a model which performs too well on the training data relative to the validation data,
Overfitting is seen when the validation loss starts to increase after reaching a global minimum. Alternately, underfitting occurs when the validation accuracy is still increasing when model optimization is terminated\mt{, because model performance is still improving with continued training}. Neither of these outcomes appears in \autoref{fig:training-accuracy}, indicating that the model optimization process was halted at an appropriate epoch.

%\svp{Starting into How CoNNOR performs}
\subsection{Model Accuracy}
